Exploring United States Housing Affordability

### 2016-08-09 paper/Paper.Rmd
edit spelling
Switch the position of a variable definition

  The inital plot measures the housing burden levels across the United States:
```{r, echo = F, fig.cap="Median Burden across Years and Household sizes"}
clean_years <- read_csv("../clean_data/clean_years.csv")

@@

```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=5, fig.height=5}
clean_years %>%
  group_by(year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(aplmed, na.rm = T)) %>%
  mutate(region=as.character(region)) %>% 
  ggplot(aes(x = year,   
                  y = new_y,   
                  colour = region)) + 
  geom_line(size = 1) + 
  labs(x = "Year", y = "Median Adjusted Income (per persons)") + 
  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
  scale_colour_brewer(palette = "Dark2", 
                      name = "Region", 
                      labels = regionNames) +
  theme_minimal(base_family = "serif")

clean_years %>%
  group_by(year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(ablmed, na.rm = T)) %>%
  mutate(region = as.character(region)) %>% 
  ggplot(aes(x = year,   
                  y = new_y,   
                  colour = region)) + 
  geom_line(size = 1) +
  labs(x = "Year", y = "Median Adjusted Income (per bedroom)") +
  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
  scale_colour_brewer(palette = "Dark2",
                      name = "Region", 
                      labels = regionNames) +
  theme_minimal(base_family = "serif")

@@

  The findings in this paper help elucidate income and househodl burden trends overtime. However, there are many holes in the actual data. The data grabbed uses gross incomes and groups data over large regions of the United States. While the spreads of incomes and housing costs may be low, the internal varation within each region will be fairly high. As a result, this does not help protray a full picture of burden level across the United States, namely in (sub)urban versus non urban areas. Also, the usage of gross income is inaccurate, as the median net income levels varied over time (even after adjusting for inflation), since taxation levels were much higher in the 1950s and 1960s than they are now- which means the burden levels are likely undervalued in the data during those decades.
  Further analysis could involve predicting the trends for the burden and income data to corroborate whether or not a two year lag exists between the Northeast and the other regions following the financial meltdown. Also, adjusting income by cost of living would provide a better understanding of burden amongst differing regions. Regions could be further subsetted by town or district level to provide more granular data.

>>>>

  The inital plot measures the housing burden levels across the United States:

```{r, echo = F, fig.cap="Median Burden across Years and Household sizes"}
clean_years <- read_csv("../clean_data/clean_sample.csv")

@@

```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=4, fig.height=4}
clean_years %>%
  group_by(year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(aplmed, na.rm = T)) %>%
  mutate(region=as.character(region)) %>% 
  ggplot(aes(x = year,   
                  y = new_y,   
                  colour = region)) + 
  geom_line(size = 1) + 
  labs(x = "Year", title = "Median Adjusted Income (per persons)") + 
  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
  scale_colour_brewer(palette = "Dark2", 
                      name = "Region", 
                      labels = regionNames) +
  theme_minimal(base_family = "serif")

clean_years %>%
  group_by(year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(ablmed, na.rm = T)) %>%
  mutate(region = as.character(region)) %>% 
  ggplot(aes(x = year,   
                  y = new_y,   
                  colour = region)) + 
  geom_line(size = 1) +
  labs(x = "Year", title = "Median Adjusted Income (per bedroom)") +
  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
  scale_colour_brewer(palette = "Dark2",
                      name = "Region", 
                      labels = regionNames) +
  theme_minimal(base_family = "serif")

@@

  The findings in this paper help elucidate income and household burden trends overtime. However, there are many holes in the actual data. The data grabbed uses gross incomes and groups data over large regions of the United States. While the spreads of incomes and housing costs may be low, the internal varation within each region will be fairly high. As a result, this does not help protray a full picture of burden level across the United States, namely in (sub)urban versus non urban areas. Also, the usage of gross income is inaccurate, as the median net income levels varied over time (even after adjusting for inflation), since taxation levels were much higher in the 1950s and 1960s than they are now- which means the burden levels are likely undervalued in the data during those decades.
  Further analysis could involve predicting the trends for the burden and income data to corroborate whether or not a two year lag exists between the Northeast and the other regions following the financial meltdown. Also, adjusting income by cost of living would provide a better understanding of burden amongst differing regions. Regions could be further subsetted by town or district level to provide more granular data.
